Numerical simulations of the Cordilleran ice sheet through the last glacial cycle

### 2014-07-17 cordillera-cycle.tex
Polish section 4.2.1.

\todo{Put more equal weight to the northern sector}

\todo{Add some point, refer back to Fig.~\ref{fig:atm}.}

\todo{Mention the major nucleation centres. Growth of the CIS from the
      highlands.}

Although snapshots from the model output (Fig.~\ref{fig:snapshots}) yield a
clear spatial picture of modelled glaciation patterns at specific time frames,
they often can not be directly compared to geomorphological evidence, which is
by nature time-transgressive. In this section, we use high-resolution
simulations driven by the GRIP and EPICA records to locate the major nucleation
centres of the Cordilleran ice sheet and assess their longevity during the last
glacial cycle.

\todo{Rewrite this opening paragraph. Fig.~\ref{fig:snapshots} can actually be
      compared to geological evidence.}

During most of the glacial cycle, modelled ice cover is restricted to disjoint
ice caps centred on major mountain ranges of the North American Cordillera
(Fig.~\ref{fig:duration}). A 2500\,km-long continuous ice cover spanning from
the Alaska Range in the northeast to the Rocky Mountains in the southwest only
exists\renote{
    I chosed to write about model output in the present form to discern in from
    the reality.}
for at most 29\,kyr in total over the entire last glacial cycle. However,
except for its marine margin and the northern foothills of the Alaska Range,
the maximum extent of the ice sheet is attained for an even more brief period
of a few thousand years (Figs.~\ref{fig:timeseries}--{fig:duration}). This
result illustrates that the maximum extents of the modelled ice sheet during
MIS~4 and 2 were both short-lived and therefore out of balance with
contemporary climate, which confirms previous inferences of a short-lived
Cordilleran ice sheet based on the geological record \citep{Clague.etal.1980,
Stroeven.etal.2010}.

>>>>

The accuracy of palaeo-glaciological reconstructions is generally biased
towards the stages of maximum extent, for which the geological evidence is
generally more detailed and more complete than for other periods. However,
these maximum stages are, by nature, extreme configurations, which do not
necessarily represent the dominant patterns of glaciation throughout the period
of ice cover.

For the Cordilleran ice sheet, geological evidence from radiocarbon dating
    \citep{Clague.etal.1980, Clague.1985, Clague.1986, Porter.Swanson.1998,
           Menounos.etal.2008},
cosmogenic exposure dating
    \citep{Stroeven.etal.2010, Stroeven.etal.2014, Margold.etal.2014},
bedrock deformation in response to former ice loads
    \citep{Clague.James.2002, Clague.etal.2005},
and offshore sedimentary records
    \citep{Cosma.etal.2008, Davies.etal.2011}
indicate that the LGM maximum extent was short-lived. To compare this finding
to our simulations, we use numerical modelling output covering the whole
simulation period to compute durations of ice cover throughout the last glacial
cycle (Fig.~\ref{fig:duration}).

The resulting maps show that, during most of the glacial cycle, modelled ice
cover is restricted to disjoint ice caps centred on major mountain ranges of
the North American Cordillera (Fig.~\ref{fig:duration}, blue areas). A
2500\,km-long continuous ice cover spanning from the Alaska Range in the
northeast to the Rocky Mountains in the southwest only exists\renote{
    I chosed to write about model output in the present form to discern in from
    the reality.}
for at most 29\,ka in total over the entire last glacial cycle
(Fig.~\ref{fig:duration}, hatched areas). However,
except for its marine margin and the northern foothills of the Alaska Range,
the maximum extent of the ice sheet is attained for an even more brief period
of a few thousand years (Figs.~\ref{fig:duration}, red areas). This
result illustrates that the maximum extents of the modelled ice sheet during
MIS~4 and 2 were both short-lived and therefore out of balance with
contemporary climate.

A notable exception to the transient character of the LGM Cordilleran ice sheet
is the northern slope of the Alaska Range, where modelled glaciers are confined
to the foothills during the entire simulation period (Fig.~\ref{fig:duration},
AR). This apparent insensitivity of modelled glacial extent to temperature
fluctuations results from a combination of low precipitations, high summer
temperatures and high temperature standard deviations in this sector of the
modelling domain (Fig.~\ref{fig:atm}). This result could potentially explain
the local distribution of glacial deposits, which indicates that these glaciers
have remained small throughout the Pleistocene \citep{Kaufman.Manley.2004}.
